---
title: "Network Correlation Analysis (NECorr)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Contents

-   [Overview](#overview)
-   [Repo Contents](#repo-contents)
-   [System Requirements](#system-requirements)
-   [Installation Guide](#installation-guide)
-   [Instructions for Use](#instructions-for-use)
-   [License](./LICENSE)
-   [Citation](#citation)

# Overview {#overview}

NECorr is a R package based on multiple-criteria decision-making algorithms. With the objective of ranking genes and their interactions in a selected condition or tissue, NECorr uses molecular network topology as well as global transcriptomics analysis to find condition/tissue-specific hub genes and their regulators.

# Repo Contents {#repo-contents}

-   [R](./R): `R` package code.
-   [man](./man): package manual for help in R session.
-   [int](./data): test data using dummies datasets and model.
-   [tests](./tests): `R` test with the data.

# System Requirements {#system-requirements}

## Hardware Requirements

The runtimes below are generated using a computer with the recommended specs (16 GB RAM, 4 [cores\@3.3](mailto:cores@3.3){.email} GHz) and internet of speed 25 Mbps.

## Software Requirements

### OS Requirements

This package is supported for *Mac OS*, *Windows* and *Linux* operating systems. The package has been tested on the following systems:

Linux: Redhat RHEL 7.4\
Mac OSX: OS X 11.2.3

The NECorr packages requires the R version 3.4.2 or higher and a standard computer with enough RAM to support the operations defined by a user. For minimal performance, this will be a computer with about 1 GB of RAM.

#### Installing R version 3.4.2 on MacOs

the latest version of R can be installed as follows:

```{bash eval=FALSE, include=TRUE}
wget https://cran.rstudio.com/bin/macosx/R-3.4.2.pkg
sudo installer -pkg R-3.4.2.pkg -target /
rm R-3.2.3.pkg
```

Two minutes will be required to install R on *Mac Os*.

#### Package dependencies

Users should install the following packages prior to installing `NECorr`, from an `R` terminal:

```{bash eval=FALSE, include=TRUE}
install.packages(c("Rcpp", "RcppParallel", "klaR", "igraph", "foreach", "doSNOW", "RColorBrewer", "gplots", "devtools", "hash"))
source("https://bioconductor.org/biocLite.R")
biocLite(c("Biobase", "limma", "supraHex"))
install.packages("dnet")

```

## Package Versions

The `NECorr` package functions with all packages in their latest versions as they appear on `CRAN` on October 15, 2017. Users can check [CRAN snapshot](https://mran.microsoft.com/timemachine/) and [BioConductor](https://www.bioconductor.org/) for details. The versions of software are, specifically:

```{r eval=FALSE, include=TRUE}
Rcpp: 0.12.17
klaR: 0.6-12
igraph: 1.0.1
dnet: 1.1.3 
foreach: 1.4.3
doSNOW: 1.0.15 
RColorBrewer: 1.1-2
gplots: 3.0.1
devtools: 1.13.2
Biobase: 2.36.2
limma: 3.32.10 
hash: 2.2.6
supraHex: 1.14.0
```

# Installation Guide {#installation-guide}

From an `R` session, type:

```{r eval=FALSE, include=TRUE}
require(devtools)
install_github('warelab/NECorr', build_vignettes=TRUE, dependencies=FALSE, upgrade_dependencies=FALSE)  # install NECorr with the vignettes
```

The package should take approximately 20 seconds to install with vignettes on a recommended computer.

## Instructions for Use {#instructions-for-use}

Please see the vignettes for help using the package:

```{r eval=FALSE, include=TRUE}
vignette("Necorr", package="NECorr")
vignette("gini", package="NECorr")
```

## Demo

Use the the scripts with the small demo dataset present in the data folder:

```{r eval=FALSE, include=TRUE}
tests/test.script.R
```

# Citation {#citation}

For citing code or the paper, please use [this citation](https://www.biorxiv.org/content/early/2018/05/21/326868).
